Off the beaten path: \\ Gaia reveals GD-1 stars outside of the main stream

### 2018-05-01 paper/gd1.tex
tweaks

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold
streams of stars that are extremely valuable tracers of the large- and
small-scale distribution of mass in the Galaxy.
Using data from the \gaia\ second data release combined with
\pans\ photometry, we present a sample of highly-probable members of the
longest cold stream in the Milky Way -- GD-1.
The resulting map of GD-1: (1) extends the apparent length of the stream by
$20^\circ$, (2) reveals a plausible location for the progenitor, (3) detects 
high-contrast gaps along the stream, and (4) indicates
the existence of stream members perturbed off the main stream track.
These discoveries were only possible because the exquisite proper motions from \gaia\ permit a clean separation of the stream from the Milky Way stars.
The additional length and proper treatment of the progenitor will aid in dynamical modeling of GD-1 and map the Galaxy on large scales.
The complex morphology of the stream points to a turbulent history; the detailed phase-space of the perturbed stream members could potentially constrain dark matter substructure in the Milky Way.

@@

To study the global properties of the stream as a function of stream longitude,
we extract stream stars around a second-order polynomial model for the stream
latitude, $\phi_2$, as a function of stream longitude, $\phi_1$.
We find the stream midpoints by computing the median $\phi_2$ in $4^\circ$
windows shifted by $2^\circ$ along $\phi_1$, then fit a parabola to the median
positions to find $\phi_{2, \textrm{track}}(\phi_1)$.
We then define the stream region as $\left| \phi_2 - \phi_{2,
\textrm{track}}(\phi_1) \right| < 0.75^\circ$.
\figurename~\ref{fig:track-and-model}, panel (a), shows the high-confidence
stream stars (same as \figurename~\ref{fig:selection}), and the two curved lines
(blue) show the upper and lower boundary of the stream region.

>>>>

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold
streams of stars that are extremely valuable tracers of the large- and
small-scale distribution of mass in the Galaxy.
Using data from the \gaia\ second data release combined with
\pans\ photometry, we present a sample of highly-probable members of the
longest cold stream in the Milky Way -- GD-1.
The resulting map of GD-1: (1) extends the apparent length of the stream by
$20^\circ$, (2) reveals a plausible location for the progenitor, (3) detects 
high-contrast gaps along the stream, and (4) indicates
the existence of stream members perturbed off the main stream track.
These discoveries were only possible because the exquisite proper motions from \gaia\ permit a clean separation of the stream from the Milky Way stars.
The additional length and proper treatment of the progenitor will aid in dynamical modeling of GD-1 and map the Galaxy on large scales.
The complex morphology of the stream points to a turbulent history; detailed phase-space properties of the perturbed stream members could potentially constrain dark matter substructure in the Milky Way.

@@

To study the global properties of the stream as a function of stream longitude,
we extract stream stars around a second-order polynomial model for the stream
latitude, $\phi_2$, as a function of stream longitude, $\phi_1$.
We find the stream midpoints by computing the median $\phi_2$ in $4^\circ$
windows shifted by $2^\circ$ along $\phi_1$, then fit a parabola to the median
positions to find $\phi_{2, \textrm{track}}(\phi_1)$.
We define the stream region as $\left| \phi_2 - \phi_{2,
\textrm{track}}(\phi_1) \right| < 0.75^\circ$.
\figurename~\ref{fig:track-and-model}, panel (a), shows the high-confidence
stream stars (same as \figurename~\ref{fig:selection}), and the two curved lines
(blue) show the upper and lower boundary of the stream region.
